NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-02-06 Manuscript/manuscriptPS.tex

\begin{abstract}
  % insert abstract here
Phosphatidylserine (PS) is the most common negatively
charged lipid in eykaryotic membranes.
PS lipids interact with signaling and other proteins via
electrostatic interactions and direct binding, and induce
membrane fusion and phase separation together with calcium ions.
Molecular details of these phenomena are not well understood
because accurate models to interpret the experimental data has not
been available. Here, we collect a set of experimental NMR data which
can be used together with molecular dynamics (MD) simulations
to interpret the lipid headgroup structures and details of ion binding
in pure PS and mixed PS:PC lipid bilayers. We use the open collaboration
method to collect data from available MD simulation models of PS lipids.
None of the models reproduce the NMR data with experimental accuracy,
but the best models suggest that the carboxyl group in the serine headgroup
does not rotate freely. In line with the previous results for PC lipids,
none of the tested force fields correctly captures the cation binding affinity to lipid bilayers containing
PS lipids. In contrast to PC lipids, the response of PS headgroups
to the bound ions qualitatively differs from experiments in the tested MD
simulation models. The collected experimental dataset and simulation
results pave the way for improvement of lipid force fields to correctly
describe negatively charged membranes and their interactions with ions.
The work is performed in the NMRlipids open collaboration project (\url{nmrlipids.blogspot.fi}).
\end{abstract}

>>>>

\begin{abstract}
  % insert abstract here
Phosphatidylserine (PS) is the most common negatively
charged lipid in eukaryotic membranes.
PS lipids interact with signaling and other proteins via
electrostatic interactions and direct binding, and induce
membrane fusion and phase separation together with calcium ions.
Molecular details of these phenomena are not well understood,
because accurate models to interpret the experimental data have not
been available. Here, we collect a set of experimental NMR data that
can be used together with molecular dynamics (MD) simulations
to interpret the lipid headgroup structures and details of ion binding
in pure PS, and mixed PS:PC, lipid bilayers. We use the open collaboration
method to collect data from a wide range of available MD models of PS lipids.
None of the models reproduce the NMR data within experimental accuracy,
but the best headgroup models suggest that the carboxyl group in the serine headgroup
does not rotate freely. In line with the previous results for PC lipids,
none of the force fields correctly captures the cation binding affinity.
In contrast to PC lipids, the response of PS headgroups
to bound ions  in the tested MD simulation models
qualitatively differs from experiments. The collected experimental dataset and simulation
results pave the way for improvement of lipid force fields to correctly
describe the biologically relevant negatively charged membranes and their interactions with ions.
This work is part of the NMRlipids open collaboration project (\url{nmrlipids.blogspot.fi}).
\end{abstract}
